 Austral Spring Extratropical Southern Hemisphere zonally assymetric circulation

### 2021-05-31 analysis/paper/shceof.Rmd
Agrega referencia más actual
#42
Cambia PSA
(#43)

One typical way of describing the zonally asymmetric circulation is by the amplitude and phase of zonal waves obtained by Fourier decomposition of geopotential height at each latitude [e.g.
@vanloon1972;@trenberth1980;\textcolor{red}{Agregar alguna referencia más actual}].
This approach suggests that zonal waves 1 and 3 explained most of the variance of the tropospheric extratropical Southern Hemisphere circulation (@vanloon1972).
However, this methodology rely on the assumption that the circulation can be meaningfully understood in terms of zonal waves of constant amplitude along a latitude circle.
This is not valid for meridionally propagating waves or zonal waves with localised amplitudes.
In the case of the wave 3, for example, @trenberth1985 observed that it played a role in blocking events, but this was mostly due to increased amplitude of a longitudinally localised wave train instead of an hemispheric-scale zonal wave 3.

@@

We computed the level-dependent Southern Annular Mode (SAM) index as the leading EOF of year-round monthly geopotential height anomalies south of 20ºS at each level for the whole period [@baldwin2009].
We further split the SAM into its zonally symmetric and zonally asymmetric components (S-SAM and A-SAM indices respectively).
These indices were obtained by projecting the zonally asymmetric and zonally symmetric part of the SAM spatial pattern onto monthly geopotential height fields, as proposed by @campitelli2021.
The Pacific South American patterns (PSA1 and PSA2) were calculated as the third and fourth leading EOF of seasonal mean for 500 hPa geopotential height with all seasons together \textcolor{red}{No me quedo claro lo de "all seasons together"} following @mo2001.

>>>>

One typical way of describing the zonally asymmetric circulation is by the amplitude and phase of zonal waves obtained by Fourier decomposition of geopotential height at each latitude [e.g.
@vanloon1972;@trenberth1980;@turner2017].
This approach suggests that zonal waves 1 and 3 explained most of the variance of the tropospheric extratropical Southern Hemisphere circulation (@vanloon1972).
However, this methodology rely on the assumption that the circulation can be meaningfully understood in terms of zonal waves of constant amplitude along a latitude circle.
This is not valid for meridionally propagating waves or zonal waves with localised amplitudes.
In the case of the wave 3, for example, @trenberth1985 observed that it played a role in blocking events, but this was mostly due to increased amplitude of a longitudinally localised wave train instead of an hemispheric-scale zonal wave 3.

@@

We computed the level-dependent Southern Annular Mode (SAM) index as the leading EOF of year-round monthly geopotential height anomalies south of 20ºS at each level for the whole period [@baldwin2009].
We further split the SAM into its zonally symmetric and zonally asymmetric components (S-SAM and A-SAM indices respectively).
These indices were obtained by projecting the zonally asymmetric and zonally symmetric part of the SAM spatial pattern onto monthly geopotential height fields, as proposed by @campitelli2021.
The Pacific South American patterns (PSA1 and PSA2) were calculated as the third and fourth leading EOF of seasonal mean for 500 hPa geopotential height combining all seasons following @mo2001.
